```{r}
library(faraway)
library(ggplot2)
library(lmtest)
```
<br><h2>Initial Data Analysis</h2><br>
```{r}
summary(chicago)
boxplot(chicago[,7],main=names(chicago)[7])
par(mfrow=c(2,3))
for(i in 1:6) 
  boxplot(chicago[,i],main=names(chicago)[i]) 
```
<br>Income has much bigger numbers then other parameters. It would have greater weight in the regression model. Hence, to avoid this we standardize the data frame. <br>
<h2>Scaling</h2>
```{r}
ch = scale(chicago)
summary(ch)
```
<br><h2>Checking the liear Structure of the model</h2><br>
```{r}
ch= data.frame(ch)
```
```{r}
lmod_full <- lm(involact~., ch) 
lmod_full$rank
summary(lmod_full)

```
```{r}
termplot(lmod_full, partial.resid = T, terms=1)
termplot(lmod_full, partial.resid = T, terms=2)
termplot(lmod_full, partial.resid = T, terms=3)
termplot(lmod_full, partial.resid = T, terms=4)
termplot(lmod_full, partial.resid = T, terms=5)
termplot(lmod_full, partial.resid = T, terms=6)


```
<br>Structure Seems Linear<br>
<h2>Checking Normality</h2>
```{r}
qqnorm(residuals(lmod_full),ylab = "Residuals", main ="")
qqline(residuals(lmod_full))
```
<br> It looks we have fatter tails problem <br>
Using Shapiro Test for varification.<br>
```{r}
#Checking using Shapiro Test
  shapiro.test(residuals(lmod_full))
```
<br>P-value High, Accepting NUll hypotheses. Dist is Normal.<br>
<h2>Checking Error Variance</h2>
```{r}
plot(fitted(lmod_full),residuals(lmod_full))
abline(h=0)
```
<br>Looks constant varrience with few anaomles.<br>
<h2>Checking Collinearity</h2>
```{r}
X = model.matrix(lmod_full)[,-1]
cor(X)
```
```{r}
vif(X)
```
<br>Every Predictor is under 5. we are safe. (volcat has relatively high correlection with other predictors<br>
Building model without volact<br>
```{r}
lmod1_without_volcat = lm(involact ~  race + fire + theft + age + income, ch )
summary(lmod1_without_volcat)
```
<br>Model without volcat performs better.<br>
<br>Now, Checking this using anova. <br>
```{r}
anova(lmod1_without_volcat,lmod_full)
```
<br> H0: Beta(r) = 0 <br>
    H1: Beta(r) != 0<br>
  <br>
    High p-value accept Null Hyptoses. <br>
simpl words: volcat is not significat. <br>
<br>Cheking Summry of the model<br>
```{r}
summary(lmod1_without_volcat)
```
<br>P-value of income is high. means it insignoficant.<br>
<br>Lets try removing income.<br>
```{r}
lmod2_without_volcat_income = lm(involact ~ race + fire + theft + age, ch)
summary(lmod2_without_volcat_income)
```
<br>Removing it does not make much of the difference<br>
<br>Comparing 2 model using anova<br>
```{r}
anova(lmod2_without_volcat_income,lmod1_without_volcat)
```

<br> H0: Beta(r) = 0 <br>
    H1: Beta(r) != 0<br>
  <br>
 <br> Significane level = 5% <br>
    High p-value accept Null Hyptoses. <br>
simple words: income is not significat. <br>

<h2>Checking Unusual Observations</h2> 
<br><h3>Checking Leverage Points</h3>
```{r}
zips = row.names(ch)
hat_vals = hatvalues(lmod2_without_volcat_income)
halfnorm(hat_vals,labs = zips, ylab = "Leverages")
```
<br>Zip Code: 60607 seems to be high leverage points.<br>
<br>Cheking these observations<br>
```{r}
row1 = which(rownames(ch) == 60607)
ch[row1,] 
```
<br>We can observe high theft in this observations<br>
<br>See what's happens if we remove this overvations.<br>
```{r}
lmod3_modified_1 = lm(involact ~ race + fire + theft + age ,ch[-row1])
summary(lmod3_modified_1)

```
<br>No Effect. It might be a case where acutally high theft happend. As it is not effecting the model let it be in the model but report the case. <br>
<br>Further Investrigation (Checking through orignoal non standerdized data source)<br>
```{r}
theft = 3
boxplot(ch[,theft],main=names(ch)[theft])
```
<br> This observation theft value is far higer then other observation. it must be reported.

<h2>Checking Outliers</h2>
```{r}
plot(lmod2_without_volcat_income)
```
<br>Cooks distance doesnot show any evidence of any outlier. However one point could possible an outlier<br>
<br>Checking using Stundtized residuals<br>
```{r}
stud=rstudent(lmod2_without_volcat_income)
tt=stud[which.max(abs(stud))]
#Pvalue calculation
n = nrow(ch)
p = length(coef(lmod2_without_volcat_income))
print('Zip(observation) and P-value')
2*(1-pt(abs(tt),n-p-1))
print("Alpha2")
(alpha2 = 0.5/n)
```
<br>p-values is < aplha. point regognizse as outlier.<br>
<br>Now lets view this observation.<br>
```{r}
row1 = which(rownames(ch) == 60610)
ch[row1,] 
```
<br>It seems we have high fire value in this observation.<br>
```{r}
fire = 2
boxplot(ch[,fire],main=names(ch)[fire])
```
<br>We can see cluster of outliers<br>
lets Investigate and get these clustor<br>
```{r}
plot(involact~fire,ch,xlab="Fire", ylab = "involact")
lmod_dummy = lm(involact~fire,ch)
abline(lmod_dummy)

#Beforni_Corection
qt(1-(.05/(47*2)),47-2-1)
range(rstudent(lmod2_without_volcat_income))
```
<br>Those High outiler fire observation lie in range of -3.259645  2.813745<br>
<br>Lets Exlcude them<br>
```{r}
#lmod_dummy2 = lm(involact ~ fire, ch, subset = (fire>3.5))
```

